---
title: "regression_on_houseprice"
output: pdf_document
date: '2022-07-17'
---


## R Markdown

```{r}
library(tidyverse)
```

```{r}
dataset <- read.csv("https://raw.githubusercontent.com/macmacmacmac/ds501_casestudy/main/data.csv") %>% na.omit
```



"City" is a categorical values and won't work with our multivariate linear
regression. To amend this, I'm sorting cities in order of being the most "urban"
(defined as having the most houses) to least urban. After which I assign each
city a number based on their ranking in the sort. City is then replaced with
this numeric "urbanness" variable. Other location variables like street,
statezip, or country are also removed. Date is useless and is removed.
```{r}
city_occurences <- as.data.frame(table(dataset$city)) %>% arrange(Freq)
dataset$urbanness <- match(dataset$city, city_occurences$Var1)


dataset <- dataset[,-which(colnames(dataset) %in% c("date", "street", 
                                                    "statezip", 
                                                    "city", "country"))]
```


I then plotted a histogram of all the variables to get a better "feel" of the 
dataset and identify any variables that needs to be transformed.
```{r}
hist(
  dataset$price,
  main = "Histogram of price",
  xlab = "price range"
)
hist(
  dataset$bedrooms,
  main = "Histogram of # of bedrooms",
  xlab = "# of bedrooms range"
)
hist(
  dataset$bathrooms,
  main = "Histogram of # of bathrooms",
  xlab = "# of bathrooms range"
)
hist(
  dataset$sqft_living,
  main = "Histogram of square foot of living",
  xlab = "squre foot of living range"
)
hist(
  dataset$sqft_lot,
  main = "Histogram of square foot of lot",
  xlab = "squre foot of lot range"
)
hist(
  dataset$floors,
  main = "Histogram of # of floors",
  xlab = "# of floors range"
)
hist(
  dataset$condition,
  main = "Histogram of condition rating",
  xlab = "condition rating range"
)
hist(
  dataset$sqft_above,
  main = "Histogram of square foot above",
  xlab = "squre foot above range"
)
hist(
  dataset$sqft_basement,
  main = "Histogram of square foot of basement",
  xlab = "squre foot of basement range"
)
hist(
  dataset$yr_built,
  main = "Histogram of year built",
  xlab = "year built range"
)
hist(
  dataset$yr_renovated,
  main = "Histogram of year renovated",
  xlab = "year renovated range"
)
hist(
  sqrt(dataset$urbanness),
  main = "Histogram of urbanness",
  xlab = "urbanness range"
)
```
Most of these seem to be reasonably distributed such that no transformations are 
egregiously needed. However, two variables that stand out are the price and the
condition

For price, there are many houses in the 6 figure range, with exponentially fewer 
and fewer houses as the price goes up. For this a log transform was applied. 
Result is a much nicer bell-curve distribution of prices. 
```{r}
hist(
  log(dataset$price),
  main = "Histogram of log(price)",
  xlab = "log(price) range"
)
```
For condition, most houses sit at a rating of 3, 4, and 5, with very few houses 
in the 1 and 2 range. A better representation of this data might be to combine 
1 and 2 into one rating, representing "poor" quality houses.

```{r}
hist(
  ifelse(dataset$condition < 3, 0, dataset$condition-2),
  main = "Histogram of square foot above",
  xlab = "squre foot above range"
)
```

Also: the two most expensive houses and the least expensive house are weird 
outliers and should be omitted. Indices are 4351, 4347, 4352.

Further, there's houses with a price of 0 for some reason? Starting to think 
this may not have been the best dataset
```{r}
dataset <- dataset[c(-4351, -4347, -4352)]
dataset <- dataset %>% dplyr::filter(price > 0)
```

Now I examine their relationship with the log of price to see if there are any
obvious trends that jump out right away. 
```{r}
normalize <- function(x, ...) {
    return((x - min(x, ...)) /(max(x, ...) - min(x, ...)))
}
```

```{r}
plot(normalize(dataset$bedrooms), 
     log(dataset$price))
plot(normalize(dataset$bathrooms), 
     log(dataset$price))
plot(normalize(dataset$sqft_living), 
     log(dataset$price))
plot(normalize(dataset$sqft_basement), 
     log(dataset$price))
plot(normalize(dataset$sqft_lot), 
     log(dataset$price))
plot(normalize(dataset$sqft_above), 
     log(dataset$price))
plot(normalize(dataset$floors), 
     log(dataset$price))
plot(normalize(ifelse(dataset$condition < 3, 0, dataset$condition)), 
     log(dataset$price))
plot(normalize(dataset$yr_built), 
     log(dataset$price))
plot(normalize(dataset$urbanness), 
     log(dataset$price))
plot(normalize(dataset$waterfront), 
     log(dataset$price))
plot(normalize(dataset$view), 
     log(dataset$price))
```
From these plots, it seems the most promising regression is square foot of 
living space, # of bathrooms, and square foot above. For every other variable, 
there is almost no variation in Y in response to variation in X. 


Hence we move on to modeling the data with MLR using 
-sqft_living
-bathrooms
-sqft_above
as our predictor variables.

We begin by normalizing the data for our model and including all other relevant 
transformations.
```{r}

model.data <- as.data.frame(apply(dataset, 2, normalize))
model.data$price <- dataset$price
model.data <- model.data %>% 
  mutate(price = log(price),
         ifelse(condition < 0.6, 0, condition-0.4),
         bedrooms = -1*bedrooms*bedrooms)
```

We then subset the data into trainning/test sets. 
```{r}
model.data.train <- model.data[sample(nrow(model.data), 0.8*nrow(model.data)),]
model.data.test <- model.data[sample(nrow(model.data), 0.2*nrow(model.data)),]
```

And now we model!
```{r}
model = lm(price~ bathrooms + 
             sqft_living + 
             sqft_above,
           data = model.data.train)

summary(model)
```
Definitely not the most accurate linear regression in the world...I think this
is close to the limitation of this hypothesis subspace and for a better 
performance, I'd have to look into deep learning. There may be some 
transformations that can help improve performance, but I think it is marginal at
best.

Predictions on test set
```{r}
predictors <- data.frame("sqft_living" = 12318,
                           "bathrooms" = 2,
                           "sqft_above" = 14423)
predictions <- as.data.frame(predict(model, predictors, interval = "confidence"))

```

```{r}
realVpredictions <- data.frame(exp(model.data.test$price), 
                                  predictions)
head(realVpredictions)
```


